---
title: "indicator_species"
author: "Anders Kiledal"
date: "7/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "D:/Google Drive/Documents/UDel/Research/Concrete_analysis/")
```


(1) Component ‘A’ is the probability that the surveyed
site belongs to the target site group given the fact that the species has
been found. This conditional probability is called the specificity or positive
predictive value of the species as indicator of the site group. (2) Component
‘B’ is the probability of finding the species in sites belonging to the site group.
This second conditional probability is called the fidelity or sensitivity of the
species as indicator of the target site group.


#Indicators for ASR
```{r}
library(tidyverse)
library(qiime2R)
library(indicspecies)
library(permute)

conc_phylo <- qiime2R::qza_to_phyloseq(features = "data/processed/biom_tables/concrete_only/all_w_taxonomy_decontam_table.biom_concrete.biom_table.qza",taxonomy = "data/qiime2/all_taxonomy.qza", tmp = tempdir())

metadata <- read_delim("data/sample-metadata.tsv", "\t", escape_double = FALSE, trim_ws = TRUE) %>%
  mutate(asr_half = paste0(ASR,"_",Halves))

taxonomy <- read_delim("data/processed/silva_taxonomy.tsv", "\t", escape_double = FALSE, trim_ws = TRUE) %>% select(otu = "#OTU ID", -confidence, taxonomy) 


samples_to_keep <- which(metadata$X.SampleID != "021715U_6")

sample_data(conc_phylo) <- metadata %>% column_to_rownames("#SampleID")

raw_series_phyloseq <- conc_phylo %>% prune_samples(sample_names(.) !="021715U_6",.)

#Filter for second half of data and taxa observed at least 5 times in 5 samples
ps <- raw_series_phyloseq %>% filter_taxa(function(x) sum(x > 1) > 5, TRUE) #%>% subset_samples(Halves = 2)

otu_table <- otuTable(ps) %>% as.data.frame() %>% as.matrix() %>% t()

ASR <- metadata[match(rownames(otu_table),metadata$`#SampleID`),] %>% .$asr_half




ind_val <- multipatt(otu_table,ASR, max.order = 1, control = how(nperm = 999))
  
summary(ind_val, indvalcomp = TRUE)



inds <- ind_val$sign %>% rownames_to_column("otu") %>% filter(  p.value <= 0.1) %>% left_join(taxonomy)

# plot.taxa(inds$otu, smooth_method = "lm" )

```


##Do the same but for combinations of species

```{r}


ASR <- metadata[match(rownames(otu_table),metadata$`#SampleID`),] %>% .$ASR

best_single_indicators <- inds %>% rownames_to_column("otu") %>% filter(p.value < 0.2) %>% .$otu

combined_indicator_table <- combinespecies(otu_table[,best_single_indicators], max.order = 2, FUN = sum)$XC


multi_ind_val <- multipatt(combined_indicator_table,ASR, control = how(nperm = 999))
  
summary(multi_ind_val, indvalcomp = TRUE)

multi_inds <- multi_ind_val$sign

```


##Do the same but for combinations of species in a more efficient way

```{r}


ASR <- metadata[match(rownames(otu_table),metadata$`#SampleID`),] %>% .$asr_half


B <- strassoc(otu_table,cluster = ASR, func = "indval")
sel <- which(B[,2]>0.02)


sc <- indicators(X = otu_table[,sel], cluster =  ASR, group = "reactive_2", verbose = FALSE, At = 0.2, Bt = 0.2,min.order = 2)

ind_table <- print(sc)

summary(sc)

print(sc, sqrtIVt = 0.6)

#Plot the coverage
plotcoverage(sc)
plotcoverage(sc, max.order=1, add=TRUE, lty=2, col="red")
legend(x=0.1, y=20, legend=c("Species combinations","Species singletons"), lty=c(1,2), col=c("black","red"), bty="n")

```



#Look for indicator genera
```{r}
library(tidyverse)
library(qiime2R)
library(readr)
library(indicspecies)
library(permute)
library(phyloseq)

set.seed(300)

conc_phylo <- qiime2R::qza_to_phyloseq(features = "data/processed/biom_tables/concrete_only/all_w_taxonomy_decontam_table.biom_concrete.biom_table.qza", 
                                       taxonomy = "data/qiime2/silva_taxonomy.qza")

metadata <- read_delim("data/sample-metadata.tsv", "\t", escape_double = FALSE, trim_ws = TRUE) %>%
  mutate(asr_half = paste0(ASR,"_",Halves))

taxonomy <- read_delim("data/processed/silva_taxonomy.tsv", "\t", escape_double = FALSE, trim_ws = TRUE) %>% select(otu = "#OTU ID", -confidence, taxonomy) %>% mutate(taxonomy = str_remove(taxonomy,"; s_.*")) 

samples_to_keep <- which(metadata$X.SampleID != "021715U_6")

sample_data(conc_phylo) <- metadata %>% column_to_rownames("#SampleID")

raw_series_phyloseq <- conc_phylo %>% prune_samples(sample_names(.) !="021715U_6",.)

#If performing on only the second half, also filters for taxa seen at least 5 times in 5 samples
#genus_Phyloseq <- speedyseq::tax_glom(raw_series_phyloseq,"Genus") %>% filter_taxa(function(x) sum(x > 5) > 5, TRUE) %>% subset_samples(Halves = 2)

#Just filter for taxa observed at least 5 times in 5 samples
genus_Phyloseq <- speedyseq::tax_glom(raw_series_phyloseq,"Genus") %>% filter_taxa(function(x) sum(x > 5) > 5, TRUE)

genus_ps_tab <- otu_table(genus_Phyloseq) %>% as.data.frame() %>% rownames_to_column("otu") %>% left_join(taxonomy) %>% mutate(tax_glom = str_extract(taxonomy, ".*g__[A-z0-9\\s]*")) %>% select(-taxonomy,-otu) %>% column_to_rownames("tax_glom") %>% t()


ASR <- metadata[match(rownames(genus_ps_tab),metadata$`#SampleID`),] %>% .$ASR

genus_ind_val <- multipatt(genus_ps_tab,ASR, control = how(nperm = 999))
  
summary(genus_ind_val, indvalcomp = TRUE)

genus_inds <- genus_ind_val$sign

reactive_ind <- genus_inds %>% 
  rownames_to_column("taxonomy") %>% 
  #filter(p.value <= 0.1, stat > 0.4 , s.reactive ==1) %>% 
  filter(p.value <= 0.1, s.reactive_2 ==1) %>% 
  mutate(tax_glom = str_extract(taxonomy, "g__[A-z]*"), tax_glom = str_remove(tax_glom,"g__"))

#plot.taxa(unmit_ind$tax_glom,smooth_method = "lm", agglom = T)

```



#Compare to another method (SIMPER) that looks at which taxa drive differences in bray curtis distances between the groups
```{r}
library(vegan)
tab <- qiime2R::qza_to_phyloseq(features = "data/processed/biom_tables/raw/all_w_taxonomy_table.biom_table.qza", 
                                       taxonomy = "data/qiime2/silva_taxonomy.qza") %>%
  speedyseq::tax_glom("Genus") %>% 
  otu_table() %>% .@.Data %>% as.data.frame() %>% rownames_to_column("otu") %>% 
  left_join(taxonomy) %>% 
  mutate(tax_glom = str_extract(taxonomy, ".*g__[A-z0-9\\s]*")) %>% 
  select(-taxonomy,-otu) %>% group_by(tax_glom) %>% summarise_at(vars(-one_of("tax_glom")),~sum(.)) %>% column_to_rownames("tax_glom") %>% t()

metadata <- read_delim("data/sample-metadata.tsv", "\t", escape_double = FALSE, trim_ws = TRUE)

taxonomy <- read_delim("data/processed/silva_taxonomy.tsv", "\t", escape_double = FALSE, trim_ws = TRUE) %>% select(otu = "#OTU ID", -confidence, taxonomy) %>% mutate(taxonomy = str_remove(taxonomy,"; s_.*")) 


simp_met <- metadata[match(rownames(tab),metadata$`#SampleID`),]


mit_simper <- simper(tab,simp_met$Type, permutations = 100, parallel = 4)

simper_df <- mit_simper$Concrete_Glass %>% as.data.frame() %>% filter(p < 0.2)

```



#Indicators for series halfs
```{r}
library(tidyverse)
library(qiime2R)
library(readr)
library(indicspecies)
library(permute)
library(phyloseq)

taxonomy <- read_tsv("data/processed/silva_taxonomy.tsv") %>%
  select(otu = "#OTU ID", taxonomy)

#For decontaminated table grouped at the genus level
otu_table <- qza_to_phyloseq(features = "data/processed/biom_tables/concrete_only/all_w_taxonomy_decontam_table.biom_concrete.biom_table.qza",
                           taxonomy = "data/qiime2/silva_taxonomy.qza") %>%
          speedyseq::tax_glom("Genus") %>%
          otu_table() %>% .@.Data %>%
          t() %>% as.data.frame()



metadata <- read_delim("data/sample-metadata.tsv", "\t", escape_double = FALSE, trim_ws = TRUE)

groups <- metadata[match(rownames(otu_table),metadata$`#SampleID`),] %>% .$Halves 

ind_val <- multipatt(otu_table,groups, max.order = 1, control = how(nperm = 999))
  
summary(ind_val, indvalcomp = TRUE)

half_inds <- ind_val$sign %>% 
  as.data.frame() %>% 
  rownames_to_column("otu") %>% 
  left_join(taxonomy)

#write_tsv(decontam_material_inds,"g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/Supplements/decontam_material_ind_genera.tsv")

```



#Indicators for seasons
```{r}
library(tidyverse)
library(qiime2R)
library(readr)
library(indicspecies)
library(permute)
library(phyloseq)

taxonomy <- read_tsv("data/processed/silva_taxonomy.tsv") %>%
  select(otu = "#OTU ID", taxonomy)

#For decontaminated table grouped at the genus level
otu_table <- qza_to_phyloseq(features = "data/processed/biom_tables/concrete_only/all_w_taxonomy_decontam_table.biom_concrete.biom_table.qza",
                           taxonomy = "data/qiime2/silva_taxonomy.qza") %>%
          speedyseq::tax_glom("Genus") %>%
          otu_table() %>% .@.Data %>%
          t() %>% as.data.frame()



metadata <- read_delim("data/sample-metadata.tsv", "\t", escape_double = FALSE, trim_ws = TRUE)

groups <- metadata[match(rownames(otu_table),metadata$`#SampleID`),] %>% .$Season

ind_val <- multipatt(otu_table,groups, max.order = 1, control = how(nperm = 999))
  
summary(ind_val, indvalcomp = TRUE)

season_inds <- ind_val$sign %>% 
  as.data.frame() %>% 
  rownames_to_column("otu") %>% 
  left_join(taxonomy)

#write_tsv(decontam_material_inds,"g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/Supplements/decontam_material_ind_genera.tsv")

```





#Indicators for material type
```{r}
library(tidyverse)
library(qiime2R)
library(readr)
library(indicspecies)
library(permute)
library(phyloseq)

taxonomy <- read_tsv("data/processed/silva_taxonomy.tsv") %>%
  select(otu = "#OTU ID", taxonomy)

#For decontaminated table grouped at the genus level
otu_table <- qza_to_phyloseq(features = "data/processed/biom_tables/decontaminated/all_w_taxonomy_decontam_table.biom_table.qza",
                           taxonomy = "data/qiime2/silva_taxonomy.qza") %>%
          speedyseq::tax_glom("Genus") %>%
          otu_table() %>% .@.Data %>%
          t() %>% as.data.frame()



metadata <- read_delim("data/sample-metadata.tsv", "\t", escape_double = FALSE, trim_ws = TRUE)

groups <- metadata[match(rownames(otu_table),metadata$`#SampleID`),] %>% .$Type

ind_val <- multipatt(otu_table,groups, max.order = 3, control = how(nperm = 999))
  
summary(ind_val, indvalcomp = TRUE)

decontam_material_inds <- ind_val$sign %>% 
  as.data.frame() %>% 
  rownames_to_column("otu") %>% 
  left_join(taxonomy)

write_tsv(decontam_material_inds,"g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/Supplements/decontam_material_ind_genera.tsv")


#For decontaminated table grouped at the genus level
otu_table <- qza_to_phyloseq(features = "data/processed/biom_tables/raw/all_w_taxonomy_table.biom_table.qza",
                           taxonomy = "data/qiime2/silva_taxonomy.qza") %>%
          speedyseq::tax_glom("Genus") %>%
          otu_table() %>% .@.Data %>%
          t() %>% as.data.frame()


metadata <- read_delim("data/sample-metadata.tsv", "\t", escape_double = FALSE, trim_ws = TRUE)

groups <- metadata[match(rownames(otu_table),metadata$`#SampleID`),] %>% .$Type

ind_val <- multipatt(otu_table,groups, max.order = 3, control = how(nperm = 999))
  
summary(ind_val, indvalcomp = TRUE)

raw_material_inds <- ind_val$sign %>% 
  as.data.frame() %>% 
  rownames_to_column("otu") %>% 
  left_join(taxonomy)

write_tsv(raw_material_inds,"g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/Supplements/raw_material_ind_genera.tsv")

```

