---
title: "make_figures"
author: "Anders Kiledal"
date: "10/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = here::here())
```

This document produces all the figures used in the concrete cylinder series paper. References are made to where they were previously made, allowing more context to be see and data processing steps that are not included here.

## Figure 1: overview of concrete cylinder series

The final version of this figure is output here.

```{r}
library(patchwork)
library(tidyverse)
library(magick)
library(lubridate)
library(scales)
library(ggpubr)

metadata <- read_delim("data/sample-metadata.tsv","\t") %>%
  mutate(Collected = mdy(Collected),
         dna_yield = num_conc * 25)

(concrete_dna_conc <- metadata %>% filter(Type == "Concrete") %>%
  ggplot(aes(Collected,dna_yield, color = ASR, shape = ASR)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  theme_minimal() +
  labs(x = NULL,
       y= "DNA yield (ng)",
       color = "ASR status") +
  scale_x_date(labels=date_format("%b '%y"), 
               date_breaks = "3 months",
               limits = c(min(metadata %>% filter(Type == "Concrete") %>% pull(Collected)) - days(30), 
                          max(metadata %>% filter(Type == "Concrete") %>% pull(Collected)))) +
  scale_shape_manual(values=c(17, 21), name = "ASR") +
  scale_color_manual(values = c("darkorange2","grey40"), name = "ASR") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1), 
        legend.position = c(.2, .8),
        legend.title = element_blank())
  )

(series_temp <- metadata %>% filter(Type == "Concrete") %>%
  ggplot(aes(Collected, temp_avg)) +
  geom_point() +
  geom_line() +
  theme_minimal() +
  labs(x = NULL,
       y= "Temperature (°F)",
       color = "ASR status") +
  scale_x_date(labels=date_format("%b '%y"), 
               date_breaks = "3 months",
               limits = c(min(metadata %>% filter(Type == "Concrete") %>% pull(Collected) - days(30)), 
                          max(metadata %>% filter(Type == "Concrete") %>% pull(Collected)))) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
    lims(y = c(20,80))
    #geom_errorbarh(aes(xmin = Collected - days(30), xmax = Collected), alpha = 0.1) +
    #theme(axis.title.x=element_blank(),
    #  axis.text.x=element_blank(),
    #  axis.ticks.x=element_blank())
  )


(precursor_dna_conc <- metadata %>%
  ggplot(aes(Type,dna_yield)) +
  geom_boxplot(outlier.alpha = 0) +
  geom_jitter(width = 0.1, alpha = 0.2) +
  #scale_y_log10() +
  theme_minimal() +
  labs(x = NULL,
       y= "DNA yield (ng)",
       color = "ASR status") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
  )

# (a) is an image produced in illustrator, which must be read in
series_image <- image_read("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/other_inPrep_or_reference/Cylinder Series_embedded_updated.png")
series_summary <- grid::rasterGrob(series_image)

layout <- "
AB
CD
"

wrap_plots(series_summary, series_temp, precursor_dna_conc, concrete_dna_conc) + 
  plot_layout(design = layout) + 
  plot_annotation(tag_levels = "a") + 
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/Fig1_Cylinder_Series_overview.pdf", 
         device = cairo_pdf, width = 6, height = 3.75,dpi = 600, scale = 1.5) +
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/Fig1_Cylinder_Series_overview.eps", 
         device=cairo_ps, fallback_resolution = 600, width = 6, height = 3.75, dpi = 600, scale = 1.5) +
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/Fig1_Cylinder_Series_overview.png", 
         type = "cairo",  width = 6, height = 3.75,dpi = 600, scale = 1.5)

```

## Figure 2: Contaminant taxa composition and relative abundance

After being output here, this figure is tweaked in Illustrator to change the format of the legend.

```{r}
library(tidyverse)
library(qiime2R)
library(metacoder)

contaminants <- read_csv("data/processed/contaminants.txt", col_names = FALSE)

otu_mat <- qiime2R::read_qza("data/processed/biom_tables/concrete_only/all_w_taxonomy_table.biom_concrete.biom_table.qza")$data %>% 
              as.data.frame() %>% 
              mutate(otu_id = rownames(.)) %>% 
              filter(otu_id %in% contaminants$X1)

taxonomy <- qiime2R::read_qza("data/qiime2/silva_taxonomy.qza") %>% .$data %>% 
              as.data.frame() %>% 
              mutate(otu_id = .$Feature.ID,lineage= .$Taxon) %>% 
              select(otu_id,lineage)

met_samples <- read_delim("data/sample-metadata.tsv", "\t", escape_double = FALSE, trim_ws = TRUE) %>% 
              select(sample_id = `#SampleID`,everything()) %>% 
              filter(sample_id %in% colnames(otu_mat))

met_otus <- otu_mat %>% left_join(taxonomy) %>% 
              select(otu_id,lineage,everything()) %>% 
              filter(!is.na(lineage),lineage!="Unassigned")

#Make the MetaCodeR obj
obj <- parse_tax_data(met_otus, class_cols = "lineage", class_sep = ";",
                      class_key = c(tax_rank = "info", tax_name = "taxon_name"),
                      class_regex = "^(.+)__(.*)$")

#removing low abundance OTUs
#obj$data$tax_data <- zero_low_counts(obj, "tax_data", min_count = 100)

no_reads <- rowSums(obj$data$tax_data[, met_samples$sample_id]) == 0
sum(no_reads)

#filtering otus with no reads
obj <- filter_obs(obj, "tax_data", ! no_reads, drop_taxa = TRUE)


#summing per-taxon counts
obj$data$rel_abd <- calc_obs_props(obj, "tax_data", other_cols = T)

obj$data$tax_abund <- calc_taxon_abund(obj, "rel_abd",
                                       cols = met_samples$sample_id,
                                       groups = met_samples$Type)

#To get in final relative abundance form
obj$data$tax_abund$Concrete <- obj$data$tax_abund$Concrete / obj$data$tax_abund$Concrete[1]

#Make the heat tree
obj %>%
taxa::filter_taxa(Concrete > 0.001) %>% #filter to only plot taxa w/ rel.abund > 0.001
heat_tree(node_label = taxon_names,
          node_size = Concrete,
          node_size_range = c(0.002,0.05),
          node_label_size_range = c(.015,.025),
          node_size_axis_label = "OTU count",
          initial_layout = "reingold-tilford", layout = "davidson-harel",
          overlap_avoidance = 10,
          node_label_max = 60 ,
          node_color = Concrete,
          node_color_range = c("gray80","gray80","gray80"),
          node_color_axis_label = "Relative abundance") +
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/Fig2_contaminants_overview.pdf", 
         device = cairo_pdf, width = 12, height = 12, dpi = 600) +
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/Fig2_contaminants_overview.eps", 
         device=cairo_ps, fallback_resolution = 600, width = 12, height = 12, dpi = 600) +
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/Fig2_contaminants_overview.png", 
         type = "cairo", dpi = 600, width = 12, height = 12)

```

## Figure 3: Concrete bacterial community relative abundance

After being output here, this figure is tweaked in Illustrator to change the format of the legend.

```{r}
library(tidyverse)
library(qiime2R)
library(metacoder)

otu_mat <- qiime2R::read_qza("data/processed/biom_tables/concrete_only/all_w_taxonomy_decontam_table.biom_concrete.biom_table.qza")$data %>%
  as.data.frame() %>% 
  mutate(otu_id = rownames(.)) 

taxonomy <- qiime2R::read_qza("data/qiime2/silva_taxonomy.qza")$data %>% 
  as.data.frame() %>% 
  mutate(otu_id = .$Feature.ID,lineage= .$Taxon) %>% 
  dplyr::select(otu_id,lineage)

met_samples <- read_delim("data/sample-metadata.tsv", "\t", escape_double = FALSE, trim_ws = TRUE) %>% 
  dplyr::select(sample_id = `#SampleID`,everything()) %>% 
  filter(sample_id %in% colnames(otu_mat))


met_otus <- otu_mat %>% left_join(taxonomy) %>% dplyr::select(otu_id,lineage,one_of(met_samples$sample_id)) %>% 
              filter(!is.na(lineage),lineage!="Unassigned")

#Make the MetaCodeR obj
  ##Greengenes class_regex = "^(.+)__(.*)$"
  ##SILVA class_regex = "^D_(.+)__(.*)$"  

obj <- parse_tax_data(met_otus, class_cols = "lineage", class_sep = ";",
                      class_key = c(tax_rank = "info", tax_name = "taxon_name"),
                      class_regex = "^D_(.+)__(.*)$")

#removing low abundance OTUs
#obj$data$tax_data <- zero_low_counts(obj, "tax_data", min_count = 300)

no_reads <- rowSums(obj$data$tax_data[, met_samples$sample_id]) == 0
sum(no_reads)

#filtering otus with no reads
obj <- filter_obs(obj, "tax_data", ! no_reads, drop_taxa = TRUE)

## Calculate relative abundance
obj$data$rel_abd <- calc_obs_props(obj, "tax_data", other_cols = T)

#summing per-taxon counts
obj$data$tax_abund <- calc_taxon_abund(obj, "rel_abd",
                                       cols = met_samples$sample_id,
                                       groups = met_samples$Type)

#To get in final relative abundance form
obj$data$tax_abund$Concrete <- obj$data$tax_abund$Concrete / obj$data$tax_abund$Concrete[1]

#Make the heat tree
obj %>%
taxa::filter_taxa(Concrete > 0.001) %>%
heat_tree(node_label = taxon_names,
          node_size = Concrete,
          node_size_trans = "area",
          node_size_range = c(0.002,0.05),
          node_label_size_range = c(.015,.025),
          node_size_axis_label = "OTU count",
          initial_layout = "reingold-tilford", 
          layout = "davidson-harel",
          #layout = "automatic",
          overlap_avoidance = 2,
          repel_force = 10,
          node_label_max = 80 ,
          node_color = Concrete,
          node_color_range = c("gray80","gray80","gray80"),
          node_color_axis_label = "Relative abundance") +
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/Fig3_concrete_community_decontam_overview.pdf", 
         device = cairo_pdf, width = 12, height = 12, dpi = 600) +
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/Fig3_concrete_community_decontam_overview.eps", 
         device=cairo_ps, fallback_resolution = 600, width = 12, height = 12, dpi = 600) +
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/Fig3_concrete_community_decontam_overview.png", 
         type = "cairo", dpi = 600, width = 12, height = 12)
```


## Figure 4: Within-sample diversity over time.

Linear mixed effects model of alpha diversity data, for DECONTAMINATED data

```{r}
library(lme4)
library(lmerTest)
library(MuMIn)
library(gridExtra)
library(grid)
library(gtable)
library(tidyverse)
library(lubridate)
library(scales)
library(patchwork)

#Load data
metadata <- read_tsv("data/sample-metadata.tsv") %>% rename(sample = "#SampleID")

shannon_table <- qiime2R::read_qza("data/processed/biom_tables/decontaminated/all_w_taxonomy_decontam_table.biom-core-metrics-results/shannon_vector.qza")$data %>% 
  rownames_to_column("sample")

faith_pd_table <- qiime2R::read_qza("data/processed/biom_tables/decontaminated/all_w_taxonomy_decontam_table.biom-core-metrics-results/faith_pd_vector.qza")$data %>% 
  rownames_to_column("sample")

separated <-  metadata %>% 
  left_join(shannon_table) %>% 
  left_join(faith_pd_table) %>% 
  select(Type, Type2,Collected,shannon,faith_pd, Months, temp_avg, ASR)

alpha_data <- separated %>% 
  gather("metric","value",-Type2,-Collected, -Months, -temp_avg, -ASR, -Type) %>% 
  mutate(Collected = lubridate::mdy(Collected))

####################
##Shannon diversity
###################

shannon_dat <- alpha_data %>% 
  filter(str_detect(Type2,"Concrete")) %>% 
  filter(metric == "shannon") %>%
  mutate(Months = scale(Months), 
         Temp = scale(temp_avg), 
         ASR = factor(ASR, levels = c("unreactive", "reactive")))


shannon_lme <- lmer(value ~ Months*ASR + Temp + (1|ASR), shannon_dat)

summary(shannon_lme)
shan_r2 <- r.squaredGLMM(shannon_lme) %>% signif(3) 
shan_r2 <- paste0("Marginal R2: ", shan_r2[1],"; Conditional R2: ",shan_r2[2])

shannon_coef <- summary(shannon_lme)$coefficients

shannon_w_pred <- modelr::add_predictions(shannon_dat,shannon_lme)


(shannon <- shannon_w_pred %>%
  select(Months, Collected, value, ASR) %>% 
  distinct() %>%
  ggplot(aes(Collected,value, shape = ASR)) + theme_bw() +
    geom_point(alpha = 0.25, size = 2) +
    geom_point(data = shannon_w_pred, aes(Collected, pred), size = 3, alpha = 0.6) +
    geom_line(data = shannon_w_pred, aes(Collected, pred, linetype = ASR), size = 1) +
    annotation_custom(textGrob(shan_r2, 0.05, 0.95,hjust = 0,gp = gpar(fontsize = 8))) +
    scale_x_date(labels=date_format("%b %y"),date_breaks = "2 months") +
    theme(axis.text.x = element_text(angle=45,vjust = 1, hjust = 1)) +
    labs(y = "Shannon", x = NULL) + 
    theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
    lims(y = c(2,8)))

shan_title <- textGrob("Shannon diversity",gp = gpar(fontsize = 16))

shannon_tab <- tableGrob(summary(shannon_lme)$coefficients %>% 
                         as.data.frame() %>% 
                         select(Estimate, p.value = `Pr(>|t|)`) %>% 
                         rownames_to_column("Coefficient") %>% 
                         filter(Coefficient != "(Intercept)") %>% 
                         mutate_at(vars(c("Estimate", "p.value")), ~ signif(.,3)) %>% 
                         mutate(p.value = if_else(p.value < 0.05, paste0(p.value,"*"),paste(p.value)),
                                p.value = if_else(p.value < 0.01, paste0(p.value,"*"),paste(p.value)),
                                p.value = if_else(p.value < 0.001, paste0(p.value,"*"),paste(p.value))), rows = NULL, theme = ttheme_default(base_size = 9))

shannon_tab <- gtable_add_rows(shannon_tab, heights = grobHeight(shan_title) + unit(5,"mm"), pos = 0)
shannon_tab <- gtable_add_grob(shannon_tab,shan_title,1,1,1,ncol(shannon_tab))

############
###Faith PD
###########

faith_dat <- alpha_data %>% 
  filter(str_detect(Type2,"Concrete")) %>% 
  filter(metric == "faith_pd") %>%
  mutate(Months = scale(Months), 
         Temp = scale(temp_avg), 
         ASR = factor(ASR, levels = c("unreactive", "reactive"))) %>% 
  distinct()


faith_lme <- lmer(value ~ Months*ASR + Temp + (1|ASR), faith_dat)

summary(faith_lme)
faith_r2 <- r.squaredGLMM(faith_lme) %>% signif(3) 
faith_r2 <- paste0("Marginal R2: ", faith_r2[1],"; Conditional R2: ",faith_r2[2])

faith_coef <- summary(faith_lme)$coefficients

faith_w_pred <- modelr::add_predictions(faith_dat,faith_lme) %>% distinct()


(faith <- faith_w_pred %>%
  select(Months, Collected, value, ASR) %>% 
  distinct() %>%
  ggplot(aes(Collected,value, shape = ASR)) + theme_bw() +
    geom_point(alpha = 0.25, size = 2) +
    geom_point(data = faith_w_pred, aes(Collected, pred), size = 3, alpha = 0.6) +
    geom_line(data = faith_w_pred, aes(Collected, pred, linetype = ASR), size = 1) +
    annotation_custom(textGrob(faith_r2, 0.05, 0.95,hjust = 0,gp = gpar(fontsize = 8))) +
    scale_x_date(labels=date_format("%b %y"),date_breaks = "2 months") +
    theme(axis.text.x = element_text(angle=45,vjust = 1, hjust = 1)) +
    labs(y = "Faith's PD", x = NULL) +
    lims(y = c(0,90)))

faith_title <- textGrob("Faith's PD",gp = gpar(fontsize = 16))

faith_tab <- tableGrob(summary(faith_lme)$coefficients %>% 
                         as.data.frame() %>% 
                         select(Estimate, p.value = `Pr(>|t|)`) %>% 
                         rownames_to_column("Coefficient") %>% 
                         filter(Coefficient != "(Intercept)") %>% 
                         mutate_at(vars(c("Estimate", "p.value")), ~ signif(.,3)) %>% 
                         mutate(p.value = if_else(p.value < 0.05, paste0(p.value,"*"),paste(p.value)),
                                p.value = if_else(p.value < 0.01, paste0(p.value,"*"),paste(p.value)),
                                p.value = if_else(p.value < 0.001, paste0(p.value,"*"),paste(p.value))), rows = NULL,theme = ttheme_default(base_size = 9))

faith_tab <- gtable_add_rows(faith_tab, heights = grobHeight(faith_title) + unit(5,"mm"), pos = 0)
faith_tab <- gtable_add_grob(faith_tab,faith_title,1,1,1,ncol(faith_tab))




##Precursor comparison box plots
shan_precursors <- separated %>% ungroup() %>%
  #filter(Type != "Concrete") %>%
  mutate(Type2 = if_else(Type2 == "Mitigated Concrete", "ASR- concrete",Type2)) %>% 
  mutate(Type2 = if_else(Type2 == "Unmitigated Concrete", "ASR+ concrete",Type2)) %>% 
  select(Type, shannon, Type2) %>%
  distinct() %>%
  ggplot(aes(Type2,shannon)) +  
  geom_boxplot() + 
  geom_jitter(size = 2, width = 0.1, alpha = 0.5) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45,hjust = 1, vjust = 1)) +
  labs(x = NULL, y = "Shannon") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
    lims(y = c(2,8))

faith_precursors <- separated %>% ungroup() %>%
  #filter(Type != "Concrete") %>%
  mutate(Type2 = if_else(Type2 == "Mitigated Concrete", "ASR- concrete",Type2)) %>% 
  mutate(Type2 = if_else(Type2 == "Unmitigated Concrete", "ASR+ concrete",Type2)) %>% 
  select(Type, faith_pd, Type2) %>%
  distinct() %>%
  ggplot(aes(Type2,faith_pd)) + 
  geom_boxplot() + 
  geom_jitter(size = 2, width = 0.1, alpha = 0.5) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45,hjust = 1, vjust = 1)) +
  labs(x = NULL, y = "Faith's PD") + 
  lims(y = c(0,90))

#Plot with precursors
design <- "
11223
44556
"

wrap_plots(shannon_tab, shannon, shan_precursors, faith_tab, faith, faith_precursors) + 
  plot_layout(design = design, guides = "collect",heights = c(1,1,0.3)) + 
  plot_annotation(tag_levels = 'a') + 
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/Fig4_alpha_diversity_decontam.pdf", 
         device = cairo_pdf, dpi = 600, width = 6.8, height = 4, scale = 1.5) +
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/Fig4_alpha_diversity_decontam.eps", 
         device=cairo_ps, fallback_resolution = 600, dpi = 600, width = 6.8, height = 4, scale = 1.5) +
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/Fig4_alpha_diversity_decontam.png", 
         type = "cairo", dpi = 600, width = 6.8, height = 4, scale = 1.5)

```

## Figure 5: Influence of precursors on concrete bacterial communities

Some text/labels in the figure were tweaked in illustrator after export.

```{r message=FALSE, warning=FALSE}
library(ggalluvial)
library(patchwork)
library(tidyverse)
library(ggtext)

#Load sourcetracking results
mit_source_prop <- read_delim("analysis/source_tracker/st_out_mit/mixing_proportions.txt", delim = "\t")
unmit_source_prop <- read_delim("analysis/source_tracker/st_out_unmit/mixing_proportions.txt", delim = "\t")

st_results <- full_join(mit_source_prop,unmit_source_prop) %>% 
  mutate(FlyAsh = if_else(is.na(FlyAsh),0,FlyAsh)) %>% 
  rename(Water = "tapwater") %>% 
  column_to_rownames("#SampleID")

#load sample metadata
metadata <- read.table('data/sample-metadata.tsv',sep='\t',h=T,check=F,comment='') %>%
  mutate(SourceSink = if_else(Type == "Concrete", "sink","source"), 
         Env= Description,) %>% 
  column_to_rownames("#SampleID")

#Load concrete composition
concrete_composition <- read_csv("analysis/source_tracker/actual_concrete_composition.csv")
```

Plot leave-one-out validation

### SF5: Sourcetracking leave one out validation

```{r}
#read in results
loo_results <- read_delim("analysis/source_tracker/st_out_mit/loo/mixing_proportions.txt", delim = "\t") %>%
  rename(sink = "collapse_col") %>%
  gather("source","val", -sink) %>%
  mutate(val = round(val, 3))

#Plot and save
(loo_heatmap <- loo_results %>% ggplot(aes(source, sink, fill = val)) +
  geom_tile() +
  geom_text(aes(label = val), color = "black") +
  coord_fixed() +
  scale_fill_gradient(low = "grey100", high = "grey50") +
  theme_minimal() +
  theme(legend.position = "none", axis.text.x = element_text(angle = 30)) + 
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/SF5_sourcetracking_loo.png", type = "cairo"))
```

```{r}
#With unknowns

for_alluvial <- st_results %>% 
  rownames_to_column("sample") %>% 
  gather("source","value", -sample) %>% 
  left_join(metadata %>% 
    rownames_to_column("sample") %>% 
    select(sample, Months, Mitigation_status, Description)) %>%
  filter(Months == 0 | Months == 21) %>%
  group_by(Description, source) %>%
  mutate(value = sum(value) / 3, type = glue::glue("Biologically inferred<br>source composition ({Months})")) %>%
  ungroup() %>%
  select(-sample, -Description) %>%
  distinct() %>%
  bind_rows(concrete_composition %>% mutate(type = "Concrete physical<br>composition", Mitigation_status = str_to_lower(Mitigation_status))) %>%
  mutate(source = factor(source,levels = c("FlyAsh","Powder","Gravel","Sand","Water","Unknown"))) %>%
  group_by(type,Mitigation_status) %>% 
  mutate(value = value / sum(value))

months_to_plot <- for_alluvial %>% filter(!is.na(Months)) %>% pull(Months) %>% unique() %>% sort()

#####################
##Unmitigated samples
#####################

unmit_source <- for_alluvial %>% filter(Mitigation_status =="unmitigated")

unmit_unknown <- unmit_source %>% ggplot(aes(factor(type, levels = c("Concrete physical<br>composition",glue::glue("Biologically inferred<br>source composition ({months_to_plot})"))), stratum = source, alluvium = source, y= value, fill = source, label = source)) +
  scale_x_discrete(expand = c(.1, .1)) +
  geom_alluvium() +
  geom_stratum(alpha = .5) +
  geom_text(stat = "stratum", size = 3) + 
  theme_minimal() + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        legend.position = "none",
        axis.text.x = element_markdown(angle = 45, hjust = 1, vjust = 1, halign = 0.5, valign = 1)) + 
  labs(y = NULL, x = NULL) + 
  ggtitle("ASR-reactive") +
  scale_fill_brewer(palette = "Greys")

###################
##Mitigated samples
###################
unmit_source <- for_alluvial %>% filter(Mitigation_status =="mitigated")

mit_unknown <- unmit_source %>% ggplot(aes(factor(type, levels = c("Concrete physical<br>composition",glue::glue("Biologically inferred<br>source composition ({months_to_plot})"))), stratum = source, alluvium = source, y= value, fill = source, label = source)) +
  scale_x_discrete(expand = c(.1, .1)) +
  geom_alluvium() +
  geom_stratum(alpha = .5) +
  geom_text(stat = "stratum", size = 3) + theme_minimal() + 
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        axis.text.y = element_blank(),
        legend.position = "none",
        axis.text.x = element_markdown(angle = 45, hjust = 1, vjust = 1, halign = 0.5, valign = 1)) + 
  labs(y = NULL, x = NULL) + 
  ggtitle("ASR-mitigated") +
  scale_fill_brewer(palette = "Greys")

#Export combined figure
unmit_unknown + mit_unknown + plot_layout(guides = 'collect') + plot_annotation(tag_levels = "a") + 
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/Fig5_SourceTracker_precursors.pdf", 
         device = cairo_pdf, dpi = 600, width = 6, height = 4) +
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/Fig5_SourceTracker_precursors.eps", 
         device=cairo_ps, fallback_resolution = 600, dpi = 600, width = 6, height = 4) +
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/Fig5_SourceTracker_precursors.png", 
         type = "cairo", dpi = 600, width = 6, height = 4)
```


## Table 1: PERMANOVA results

```{r message=FALSE, warning=FALSE}
library(vegan)

gen_unidist <- read_qza("data/processed/biom_tables/concrete_only/all_w_taxonomy_decontam_table.biom_concrete.biom-core-metrics-results/generalized_05_unifrac_distance_matrix.qza") %>% .$data

metadata <- read_delim("data/sample-metadata.tsv", "\t", escape_double = FALSE, trim_ws = TRUE) %>% 
  as.data.frame() %>% 
  filter(Type == "Concrete") %>% 
  mutate(Temperature = scale(temp_avg),
         Months = scale(Months)) %>% 
  column_to_rownames("#SampleID") %>%
  .[rownames(as.matrix(gen_unidist)),]

perm <- how(nperm = 999)
#setBlocks(perm) <- with(metadata, ASR)

set.seed(333)

##############
# sequentially
by <- "terms"
(adonis_sequential <- adonis2(gen_unidist ~ Temperature + Months + ASR, data = metadata, permutations = perm, by = by))
seq_summary <- broom::tidy(adonis_sequential) %>% mutate(terms_tested = "Sequentially")

############
# Marginally
by <- "margin"
(adonis_margin <- adonis2(gen_unidist ~ Temperature + Months + ASR, data = metadata, permutations = perm, by = by))
margin_summary <- broom::tidy(adonis_margin) %>% mutate(terms_tested = "Marginally")

#Calculate beta dispersion p.value for each term
beta_dist <- function(dist, term){
  anova(betadisper(dist,metadata %>% pull(paste0(term))))$`Pr(>F)`[1]
}

terms <- c("Temperature","Months","ASR")

disp <- data.frame(disp_p_val = sapply(terms,beta_dist,dist = gen_unidist)) %>% 
  rownames_to_column("Coefficient")


#Make summary table

permanova_table <- bind_rows(seq_summary,margin_summary) %>% 
  mutate(Metric = "gUniFrac") %>%
  filter(term %in% terms) %>% 
  select(Metric, terms_tested, Coefficient = "term", R2, F = "statistic", p.value) %>% 
  left_join(disp) %>% 
  mutate(across(c("F","p.value","disp_p_val","R2"),~round(.,3))) %>% 
  rename(`Terms Tested` = "terms_tested", `Dispersion p.value` = "disp_p_val")

knitr::kable(permanova_table)

```

## Figure 6: Taxa changes over time

Blue (both) added to legend with illustrator after output here.

```{r}
library(scales)
library(tidyverse)
library(qiime2R)
library(readr)
library(lubridate)
library(microbiome)
library(patchwork)

#Import and process (make relative abund)

otu_table <- read_qza("data/processed/biom_tables/concrete_only/all_w_taxonomy_decontam_table.biom_concrete.biom_table.qza")$data %>% as.data.frame()

metadata <- sample_metadata <- read_delim("data/sample-metadata.tsv", "\t", escape_double = FALSE, trim_ws = TRUE) %>% 
  mutate(sample = `#SampleID`, Collected = mdy(Collected))

make_rel <- function(x) {x/sum(x)}

rel_otu_table <- otu_table %>% 
  transform("compositional") %>% as.data.frame() %>% 
  mutate(otu = rownames(otu_table))

taxonomy <- read_delim("data/processed/silva_taxonomy.tsv", "\t", escape_double = FALSE, trim_ws = TRUE) %>% 
  separate(., taxonomy, c("Kingdom","Phylum","Class","Order", "Family", "Genus","Species"), sep= "; [a-z]__", remove=TRUE) %>% 
  select(-confidence, otu = `#OTU ID`) %>% 
  filter(otu %in% rownames(otu_table))

long_otu <- rel_otu_table %>% 
  gather(key = "sample", value = "abund", -otu) %>% 
  left_join(taxonomy) %>% 
  left_join(metadata[,c("sample","Months","ASR","temp_avg", "Collected")])


#Set the taxonomic level being explored
tax_level <- quo(Phylum)

#Get the most abundant taxa at the defined level
top_10_phyla <- long_otu %>% 
  group_by(!!tax_level) %>% 
  mutate(rel_sum = sum(abund)) %>% 
  select(!!tax_level, rel_sum) %>% 
  ungroup() %>% unique() %>% 
  top_n(5, rel_sum)

#Calculate the number of ASVs for each taxa
Phyla_subs <- long_otu %>% 
  filter(abund > 0) %>% 
  group_by(sample, !!tax_level) %>% 
  mutate(sub_species = n()) %>%
  ungroup() %>% group_by(!!tax_level) %>%
  mutate(subspecies.z = scale(sub_species))

#make table for plotting
Phyla_subs <- long_otu %>% 
  filter(abund > 0) %>% 
  select(otu, Phylum, Class, Order, Family, Genus, Species, Months, ASR, Collected) %>%
  distinct() %>% 
  group_by(Months, ASR, !!tax_level) %>% 
  mutate(sub_species = n()) %>%
  ungroup() %>% group_by(!!tax_level) %>%
  mutate(subspecies.z = scale(sub_species))

#Richness plot

(subspecies_richness <- Phyla_subs %>% filter(!!tax_level %in% pull(top_10_phyla, !!tax_level)) %>% 
  select(Months, subspecies.z, sub_species, !!tax_level, ASR, Collected) %>% distinct() %>% 
  group_by(!!tax_level) %>% mutate(phy_mean = mean(sub_species)) %>% 
  ggplot(aes(Collected, sub_species, linetype = ASR, shape = ASR, color = ASR)) + 
    geom_point(alpha = 0.7, size = 2) + 
    scale_shape_manual(values=c(17, 21), name = "ASR") +
    scale_color_manual(values = c("darkorange2","grey40"), name = "ASR") +
    geom_smooth(method = "loess", se = F) +
    scale_x_date(labels=date_format("%b '%y"),
                 date_breaks = paste(3,"months"),
                 limits = c(min(mdy(Phyla_subs$Collected)),max(mdy(Phyla_subs$Collected)))) +
    #geom_smooth(method = "lm", se = F) +
    scale_y_log10() + 
    labs(y = "ASV richness (log)") +
    #facet_wrap(vars( factor(!!tax_level,top_group_f)),nrow = 1) + #use this if trying to match the facet order of Shannon plot
    facet_wrap(vars( factor(!!tax_level,pull(top_10_phyla,!!tax_level))),nrow = 1) +
    theme_bw() +
    theme(axis.text.x = element_text(angle=45,vjust = 1, hjust = 1),
          axis.title.x = element_blank()))

#Prevalence plot

(prev_examples <- plot_prevalence("Acinetobacter", se = F) +
    plot_prevalence("Bacillus", se = F) +  theme(axis.text.y=element_blank(),axis.title.y=element_blank()) +  
    plot_prevalence("Beijerinckiaceae", se = F) + theme(axis.text.y=element_blank(),axis.title.y=element_blank()) + 
    plot_prevalence("Psychrobacter", se = F) + theme(axis.text.y=element_blank(),axis.title.y=element_blank()) + 
    plot_prevalence("Ferruginibacter", se = F) + theme(axis.text.y=element_blank(),axis.title.y=element_blank()) + 
    plot_layout(ncol = 5,  tag_level = "new"))

#Relative abundance plots

(rel_abund <- plot.taxa("Acinetobacter", agglom = TRUE, point_alpha = 0.7, se = FALSE) +
    plot.taxa("Bacillus", agglom = TRUE, point_alpha = 0.7, se = FALSE) + labs(y = NULL) +
    plot.taxa("Beijerinckiaceae", agglom = TRUE, point_alpha = 0.7, se = FALSE) + labs(y = NULL) +
    plot.taxa("Psychrobacter",agglom = TRUE, point_alpha = 0.7, se = FALSE) + labs(y = NULL) +
    #plot.taxa("Romboutsia",agglom = TRUE, point_alpha = 0.8, smooth_method = "lm") + labs(y = NULL) +
    plot.taxa("Ferruginibacter",smooth_method = NA, agglom = TRUE, point_alpha = 0.7, se = FALSE) + labs(y = NULL) + 
    plot_layout(ncol = 5, tag_level = "new"))


subspecies_richness / prev_examples / rel_abund + plot_layout(guides = "collect") +
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/Fig6_taxa_changes.pdf", 
         device = cairo_pdf, dpi = 600, scale = 1.5, width = 6.8, height = 4) +
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/Fig6_taxa_changes.eps", 
         device=cairo_ps, fallback_resolution = 600, dpi = 600, scale = 1.5, width = 6.8, height = 4) +
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/Fig6_taxa_changes.png", 
         type = "cairo", dpi = 600, scale = 1.5, width = 6.8, height = 4)

```

## Figure 7: Concrete compared to EMP samples

Plot the PCoA produced with QIIME 2

```{r}
library(tidyverse)
library(qiime2R)
library(glue)
library(patchwork)
library(ragg)
library(ggtext)

emp_tapwater_concrete_map <- read_delim("data/processed/emp_data/emp_tapwater_concrete_map.tsv", "\t", escape_double = FALSE, trim_ws = TRUE) %>% 
  dplyr::rename(sample = "#SampleID") %>% 
  filter(include_in_concrete_comparison == TRUE)

pcoa <- read_qza("data/processed/emp_data/core-metrics-results_emp_decontam_concrete_tapwater/general_0_5_unifrac_pcoa.qza")$data

series_samples <- c("Concrete", "FlyAsh", "Powder", "Sand", "Gravel")

empo3 <- levels(as.factor(emp_tapwater_concrete_map$empo_3)) %>% 
  data.frame(empo_3 = .) %>% 
  cbind( colors = c("grey80", "#ff0000", "#f27304", "#f27304", "#ff0000", "#ff0000", "black", "#fcc688", "maroon2", "maroon2", "#52ff3c", "#500261", "#80c99b", "#045b04", "#80c99b", "maroon2", "maroon2",  "#ff0000", "#a36f23", "#765f3d", "#6b440b", "#ffff00", "#ababab", "#5e5e5e", "#7cecf4", "#0000ff","#000098")) %>%
  mutate(shapes = case_when(empo_3 == "Powder" ~ 16,
                            empo_3 == "Gravel" ~ 15,
                            empo_3 == "FlyAsh" ~ 17,
                            empo_3 == "Sand" ~ 18, 
                            empo_3 == "Concrete" ~ 16)) %>% 
  mutate(shapes = if_else(is.na(shapes), 16, shapes),
         empo_3 = as.character(empo_3), colors = as.character(colors),
         empo_3 = if_else(empo_3 %in% series_samples, glue("**{empo_3}**"), empo_3)) #%>%
  #filter(!empo_3 %in% series_samples)

variance_explained <- pcoa$ProportionExplained %>%
  gather("pc", "variance_explained") %>%
  mutate(percent_variance_explained = glue("{round(variance_explained * 100,2)} %"))

long_pcoa <- pcoa$Vectors %>% 
  as.data.frame() %>%
  select(sample = "SampleID", PC1, PC2, PC3) %>%
  left_join(emp_tapwater_concrete_map %>% select(sample,empo_3)) %>% 
  mutate(our_sample = if_else(empo_3 %in% series_samples,TRUE, FALSE) ,
         empo_3 = if_else(empo_3 %in% series_samples, glue("**{empo_3}**"), empo_3))
  
  
pc_1_2 <- long_pcoa %>% filter(!empo_3 %in% series_samples) %>% ggplot(aes(PC1, PC2, color = factor(empo_3, levels = empo3$empo_3), label2 = sample, shape = factor(empo_3, levels = empo3$empo_3))) + 
  geom_point(size = 0.4, alpha = 0.4) +
  geom_point(data = long_pcoa %>% filter(our_sample == TRUE), size = 2.5, alpha = 0.8) +
  theme_bw() +
  coord_fixed(ratio = variance_explained$variance_explained[2] / variance_explained$variance_explained[1]) +
  scale_color_manual(values = empo3$colors) +
  scale_shape_manual(values = empo3$shapes) +
  labs(x = NULL,
       y = glue("PC2 [{variance_explained$percent_variance_explained[2]}]")) + 
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(), 
        legend.position = "none") + 
  guides(size = FALSE) 
  

pc_1_3 <- long_pcoa %>% filter(!empo_3 %in% series_samples) %>% ggplot(aes(PC1, PC3, color = factor(empo_3, levels = empo3$empo_3), label2 = sample, shape = factor(empo_3, levels = empo3$empo_3))) + 
  geom_point(size = 0.4, alpha = 0.4) +
  geom_point(data = long_pcoa %>% filter(our_sample == TRUE), size = 2.5, alpha = 0.8) +
  theme_bw() +
  coord_fixed(ratio = variance_explained$variance_explained[3] / variance_explained$variance_explained[1]) +
  scale_color_manual(values = empo3$colors) +
  scale_shape_manual(values = empo3$shapes) +
  labs(x = glue("PC1 [{variance_explained$percent_variance_explained[1]}]"),
       y = glue("PC3 [{variance_explained$percent_variance_explained[3]}]"),
       color = "Sample type (EMPO level 3)",
       shape = "Sample type (EMPO level 3)") +
  guides(size = FALSE) +
  theme(legend.text = element_markdown(size = 8))

(pcoa_plot_no_title <- pc_1_2 / pc_1_3 + plot_layout(guides = "collect") + 
    ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/other_inPrep_or_reference/EMP_PCoA_generalised_0.5_UniFrac_no_title.png", type = "cairo", width = 6.8, height = 3.8, scale = 1.2))
```

Plot mean distances of concrete samples to precursors and EMPO3 groups

```{r EMP distance comparisons}
library(qiime2R)
library(tidyverse)
library(patchwork)
library(ggtext)
library(glue)

metadata <- read_delim("data/processed/emp_data/emp_and_tapwater_comparison_map.tsv", "\t", escape_double = FALSE, trim_ws = TRUE) %>%
  dplyr::select(sample = "#SampleID",Env,empo_2,empo_3, env_material,env_feature,env_biome,sample_scientific_name,host_common_name)

series_samples <- c("Concrete", "FlyAsh", "Powder", "Sand", "Gravel")

#Import decontaminated distances

concrete_samples <- metadata %>% filter(empo_3 == "Concrete") %>% pull(sample)

decontam_emp_dists <- read_qza("data/processed/emp_data/core-metrics-results_emp_decontam_concrete_tapwater/general_0_5_unifrac_distance_matrix.qza")$data %>% 
  as.matrix() %>% as.data.frame() %>% 
  rownames_to_column("s1") %>% 
  gather("s2","decontam_dist",-s1) %>%
  filter(s1 != s2, s1 %in% concrete_samples) %>%
  left_join(metadata %>% dplyr::rename(s1 = "sample")) %>%
  dplyr::rename(s1_env = "Env", 
                s1_empo_2 = "empo_2", 
                s1_empo_3 = "empo_3", 
                s1_mat = "env_material", 
                s1_feature = "env_feature", 
                s1_biome = "env_biome", 
                s1_sci_name = "sample_scientific_name", 
                s1_host = "host_common_name") %>%
  left_join(metadata %>% dplyr::rename(s2 = "sample")) %>% 
  dplyr::rename(s2_env = "Env", 
                s2_empo_2 = "empo_2", 
                s2_empo_3 = "empo_3",
                s2_mat = "env_material", 
                s2_feature = "env_feature", 
                s2_biome = "env_biome", 
                s2_sci_name = "sample_scientific_name", 
                s2_host = "host_common_name") %>%
  filter(s1_empo_3 == "Concrete") %>% 
  gather("dist_type","dist",c("decontam_dist"))


similarity <- decontam_emp_dists %>% 
  mutate(decontam_sim = 1 - dist) %>% 
  mutate(precursor = if_else(s2_empo_3 %in% series_samples[-1], TRUE, FALSE),
         s2_empo_3 = if_else(s2_empo_3 %in% series_samples, glue("**{s2_empo_3}**"), s2_empo_3))


mean_emp_sim <- decontam_emp_dists %>%
  group_by(dist_type,s2_empo_3) %>% 
  filter(!is.na(dist), dist_type == "decontam_dist") %>%
  select(dist,s2_empo_3,dist_type) %>% distinct() %>%
  mutate(mean_dist = 1- mean(dist), mean_sd = sd(dist), sem = sd(dist)/sqrt(length(dist))) %>%
  select(s2_empo_3,mean_dist,dist_type,sem,mean_sd, dist) %>% 
  mutate(precursor = if_else(s2_empo_3 %in% series_samples[-1], TRUE, FALSE),
         s2_empo_3 = if_else(s2_empo_3 %in% series_samples, glue("**{s2_empo_3}**"), s2_empo_3))


#Plot mean distance between decontaminated concrete samples and empo3 groups + precursor materials
(emp_sim_horiz <- mean_emp_sim %>%
  ggplot(aes(reorder(s2_empo_3,-mean_dist),1- dist)) + 
    geom_violin() +
    geom_point(aes(x = s2_empo_3, y = mean_dist), data = . %>% select(s2_empo_3,mean_dist) %>% distinct(), inherit.aes = FALSE) + 
    theme_minimal() +
    theme(axis.text.x = element_markdown(size = 7,angle = 45,valign = 1,halign = 1,vjust = 1,hjust = 1),
          axis.title.y = element_text(size = 8)) +
    labs(x = NULL, y = "Mean similarity (1- gUniFrac distance)"))


#Combine plots for the paper

##Load the PCoA plot, easier to do from the image to keep proportions accurate and not make the scales go wonky
emp_pcoa_image <- magick::image_read("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/other_inPrep_or_reference/EMP_PCoA_generalised_0.5_UniFrac_no_title.png") %>% grid::rasterGrob()


##Define the plot layout
layout <- "
AAAAA
AAAAA
AAAAA
BBBBB
"

##Make plot without precursors included in the mean dists
##Make plot with precursors included in mean dists
wrap_plots(emp_pcoa_image, emp_sim_horiz) + 
  plot_layout(design = layout) + 
  plot_annotation(tag_levels = "a") +
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/Fig7_EMP_PCoA_generalised_0.5_UniFrac.pdf", 
         device = cairo_pdf, dpi = 600, width = 6, height = 6) +
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/Fig7_EMP_PCoA_generalised_0.5_UniFrac.eps", 
         device=cairo_ps, fallback_resolution = 600, dpi = 600, width = 6, height = 6) +
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/Fig7_EMP_PCoA_generalised_0.5_UniFrac.png", 
         type = "cairo", dpi = 600, width = 6, height = 6)

```

## Figure 8: Potential bioindicators of ASR

```{r}
library(patchwork)
library(tidyverse)

multiple <- c("Arcobacter", "Bryobacter", "Carnobacterium", "Lawsonella", "Modestobacter", "Salinicoccus", "Rhodocyclaceae", "Lawsonella", "Flavobacterium", "Rheinheimera soli") %>% unique()

indic_sp <- c("Arcobacter", "Bryobacter")
metaSeq <- c("Arcobacter", "Bryobacter", "Carnobacterium", "Lawsonella", "Modestobacter", "Salinicoccus")
logit <- c("Flavobacterium", "Lawsonella", "Rheinheimera soli", "Rhodocyclaceae")
  

symbols <- data.frame(tax = multiple) %>% 
  mutate(meta = if_else(tax %in% metaSeq, "a",""), 
         ind = if_else(tax %in% indic_sp, "b",""),
         log = if_else(tax %in% logit, "c",""),
         symbols = glue::glue("{meta}{ind}{log}")) %>% 
  rename(og_taxa = "tax")



mult <- plot.taxa(taxa = multiple, smooth_method = "lm", ncol = 3)

mult_df <- mult$data %>% 
  left_join(symbols) %>% 
  #mutate(tax_and_level = str_remove(tax_and_level, "([kpcofgs])")) %>% 
  mutate(tax_and_level = glue::glue("{og_taxa} ({symbols})"))

mult$data <- mult_df

#without this it didn't save w/ symbols correctly
mult_exp <- mult

mult_exp +
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/Fig8_ASR_indicators.pdf", 
         device = cairo_pdf, dpi = 600,width = 6.5, height =4) +
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/Fig8_ASR_indicators.eps", 
         device=cairo_ps, fallback_resolution = 600, dpi = 600, width = 6.5, height =4) +
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/Fig8_ASR_indicators.png", 
         type = "cairo", dpi = 600, width = 6.5, height =4)

```

## ST1: ASV count at each processing stage

```{r}
#Load metadata
metadata <- read_delim("data/sample-metadata.tsv", "\t") %>%
  select(sample = "#SampleID", Months_since_start = "Months", Collected, ASR, old_sample_ids)

#Define function for extracting data from qiime2 summary file
read_demux_qzv <- function(qzv_path){
  temp_folder <- tempdir() #create temp directory
  qzv_contents <- zip::zip_list(qzv_path) #list files in zip
  qzv_summary_file <- qzv_contents %>% #find summary table
    filter(str_detect(filename,"per-sample-fastq-counts.csv")) %>% #find summary table
    pull(filename)
  zip::unzip(qzv_path,qzv_summary_file,overwrite = TRUE, exdir = temp_folder) #unzip qzv to temp.dir
  summary_table <- read_delim(file.path(temp_folder,qzv_summary_file), ",") %>% #read summary table
    dplyr::rename(sample = "Sample name")
  file.remove(file.path(temp_folder,qzv_summary_file)) #remove temp.file
  return(summary_table)
}

#Get the qiime2 demux summary information

#Read in demux files
raw_demux_summary <- read_demux_qzv("data/qiime2/paired-end-demux.qzv") %>%
  rename(raw_sequence_count = "Sequence count", old_sample_ids = "sample")

qual_filt_demux_summary <- read_demux_qzv("data/qiime2/dada2_filt_paired-end-demux.qzv") %>%
  rename(post_cutadapt_sequence_count = "Sequence count")

#Add sample read counts to summary table and compute percent of reads passing quality filtering w/ CutAdapt
summary_table <- metadata %>% 
  left_join(raw_demux_summary) %>%
  left_join(qual_filt_demux_summary) %>%
  mutate(percent_reads_retained_cutadapt = (post_cutadapt_sequence_count/raw_sequence_count)*100)


# Def. function for extracting table summary information
read_table_summary <- function(qzv_path){
  temp_folder <- tempdir() #create temp directory
  qzv_contents <- zip::zip_list(qzv_path) #list files in zip
  qzv_summary_file <- qzv_contents %>% #find summary table
    filter(str_detect(filename,"sample-frequency-detail.csv")) %>% #find summary table
    pull(filename)
  zip::unzip(qzv_path,qzv_summary_file,overwrite = TRUE, exdir = temp_folder) #unzip qzv to temp.dir
  summary_table <- read_delim(file.path(temp_folder,qzv_summary_file), ",",col_names = FALSE) %>% #read summary table
    dplyr::rename(sample = "X1", seq_obs = "X2")
  file.remove(file.path(temp_folder,qzv_summary_file)) #remove temp.file
  return(summary_table)
}

#Import and add DADA2 stats
dada2_stats <- read_qza("data/qiime2/dada2_stats.qza")$data %>%
  rownames_to_column("sample") %>%
  rename_at(vars(-sample),~paste0("dada2.",.))

#Import decontamination summary
decontam_summary <- read_delim("data/processed/decontamination_stat.tsv","\t") %>%
  select(sample, post_decontam_reads,percent_read_loss,pre_decontam_otu_n = otu_n, post_decontam_otu_n,percent_otu_loss)


#Merge DADA2 stats and the decontam table summary
obs_table <- summary_table %>%
  left_join(dada2_stats) %>%
  left_join(decontam_summary) %>%
  select(-old_sample_ids) %>%
  arrange(lubridate::mdy(Collected)) %>%
  write_delim("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/ST_1_processing_summary.tsv","\t")

```

## ST2: Contaminants by detection method

Produced in the decontamination .Rmd document.


## ST3: Contaminant taxa

Produced in the decontamination .Rmd document.


## SF1: Euler plot of contaminants identified per method

Produced in the decontamination .Rmd document.

## SF2: Euler plot of contaminant sources

Produced in the decontamination .Rmd document.

## SF3: Heat tree matrix of contaminant taxa

Produced in the decontamination .Rmd document.

## SF4: Within sample diverity before decontamination

Linear mixed effects model of alpha diversity data, for RAW data

```{r}
library(lme4)
library(lmerTest)
library(MuMIn)
library(gridExtra)
library(grid)
library(gtable)
library(tidyverse)
library(lubridate)
library(scales)
library(patchwork)

#Load data
metadata <- read_tsv("data/sample-metadata.tsv") %>% rename(sample = "#SampleID")

shannon_table <- qiime2R::read_qza("data/processed/biom_tables/raw/all_w_taxonomy_table.biom-core-metrics-results/shannon_vector.qza")$data %>% 
  rownames_to_column("sample")

faith_pd_table <- qiime2R::read_qza("data/processed/biom_tables/raw/all_w_taxonomy_table.biom-core-metrics-results/faith_pd_vector.qza")$data %>% 
  rownames_to_column("sample")

separated <-  metadata %>% 
  left_join(shannon_table) %>% 
  left_join(faith_pd_table) %>% 
  select(Type, Type2,Collected,shannon,faith_pd, Months, temp_avg, ASR)

alpha_data <- separated %>% 
  gather("metric","value",-Type2,-Collected, -Months, -temp_avg, -ASR, -Type) %>% 
  mutate(Collected = lubridate::mdy(Collected))

####################
##Shannon diversity
###################

shannon_dat <- alpha_data %>% 
  filter(str_detect(Type2,"Concrete")) %>% 
  filter(metric == "shannon") %>%
  mutate(Months = scale(Months), 
         Temp = scale(temp_avg), 
         ASR = factor(ASR, levels = c("unreactive", "reactive")))


shannon_lme <- lmer(value ~ Months*ASR + Temp + (1|ASR), shannon_dat)

summary(shannon_lme)
shan_r2 <- r.squaredGLMM(shannon_lme) %>% signif(3) 
shan_r2 <- paste0("Marginal R2: ", shan_r2[1],"; Conditional R2: ",shan_r2[2])

shannon_coef <- summary(shannon_lme)$coefficients

shannon_w_pred <- modelr::add_predictions(shannon_dat,shannon_lme)


(shannon <- shannon_w_pred %>%
  select(Months, Collected, value, ASR) %>% 
  distinct() %>%
  ggplot(aes(Collected,value, shape = ASR)) + theme_bw() +
    geom_point(alpha = 0.25, size = 2) +
    geom_point(data = shannon_w_pred, aes(Collected, pred), size = 3, alpha = 0.6) +
    geom_line(data = shannon_w_pred, aes(Collected, pred, linetype = ASR), size = 1) +
    annotation_custom(textGrob(shan_r2, 0.05, 0.95,hjust = 0,gp = gpar(fontsize = 8))) +
    scale_x_date(labels=date_format("%b %y"),date_breaks = "2 months") +
    theme(axis.text.x = element_text(angle=45,vjust = 1, hjust = 1)) +
    labs(y = "Shannon", x = NULL) + 
    theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
    lims(y = c(2,8)))

shan_title <- textGrob("Shannon diversity",gp = gpar(fontsize = 16))

shannon_tab <- tableGrob(summary(shannon_lme)$coefficients %>% 
                         as.data.frame() %>% 
                         select(Estimate, p.value = `Pr(>|t|)`) %>% 
                         rownames_to_column("Coefficient") %>% 
                         filter(Coefficient != "(Intercept)") %>% 
                         mutate_at(vars(c("Estimate", "p.value")), ~ signif(.,3)) %>% 
                         mutate(p.value = if_else(p.value < 0.05, paste0(p.value,"*"),paste(p.value)),
                                p.value = if_else(p.value < 0.01, paste0(p.value,"*"),paste(p.value)),
                                p.value = if_else(p.value < 0.001, paste0(p.value,"*"),paste(p.value))), rows = NULL, theme = ttheme_default(base_size = 9))

shannon_tab <- gtable_add_rows(shannon_tab, heights = grobHeight(shan_title) + unit(5,"mm"), pos = 0)
shannon_tab <- gtable_add_grob(shannon_tab,shan_title,1,1,1,ncol(shannon_tab))

############
###Faith PD
###########

faith_dat <- alpha_data %>% 
  filter(str_detect(Type2,"Concrete")) %>% 
  filter(metric == "faith_pd") %>%
  mutate(Months = scale(Months), 
         Temp = scale(temp_avg), 
         ASR = factor(ASR, levels = c("unreactive", "reactive"))) %>% 
  distinct()


faith_lme <- lmer(value ~ Months*ASR + Temp + (1|ASR), faith_dat)

summary(faith_lme)
faith_r2 <- r.squaredGLMM(faith_lme) %>% signif(3) 
faith_r2 <- paste0("Marginal R2: ", faith_r2[1],"; Conditional R2: ",faith_r2[2])

faith_coef <- summary(faith_lme)$coefficients

faith_w_pred <- modelr::add_predictions(faith_dat,faith_lme) %>% distinct()


(faith <- faith_w_pred %>%
  select(Months, Collected, value, ASR) %>% 
  distinct() %>%
  ggplot(aes(Collected,value, shape = ASR)) + theme_bw() +
    geom_point(alpha = 0.25, size = 2) +
    geom_point(data = faith_w_pred, aes(Collected, pred), size = 3, alpha = 0.6) +
    geom_line(data = faith_w_pred, aes(Collected, pred, linetype = ASR), size = 1) +
    annotation_custom(textGrob(faith_r2, 0.05, 0.95,hjust = 0,gp = gpar(fontsize = 8))) +
    scale_x_date(labels=date_format("%b %y"),date_breaks = "2 months") +
    theme(axis.text.x = element_text(angle=45,vjust = 1, hjust = 1)) +
    labs(y = "Faith's PD", x = NULL) +
    lims(y = c(0,90)))

faith_title <- textGrob("Faith's PD",gp = gpar(fontsize = 16))

faith_tab <- tableGrob(summary(faith_lme)$coefficients %>% 
                         as.data.frame() %>% 
                         select(Estimate, p.value = `Pr(>|t|)`) %>% 
                         rownames_to_column("Coefficient") %>% 
                         filter(Coefficient != "(Intercept)") %>% 
                         mutate_at(vars(c("Estimate", "p.value")), ~ signif(.,3)) %>% 
                         mutate(p.value = if_else(p.value < 0.05, paste0(p.value,"*"),paste(p.value)),
                                p.value = if_else(p.value < 0.01, paste0(p.value,"*"),paste(p.value)),
                                p.value = if_else(p.value < 0.001, paste0(p.value,"*"),paste(p.value))), rows = NULL,theme = ttheme_default(base_size = 9))

faith_tab <- gtable_add_rows(faith_tab, heights = grobHeight(faith_title) + unit(5,"mm"), pos = 0)
faith_tab <- gtable_add_grob(faith_tab,faith_title,1,1,1,ncol(faith_tab))




##Precursor comparison box plots
shan_precursors <- separated %>% ungroup() %>%
  #filter(Type != "Concrete") %>%
  mutate(Type2 = if_else(Type2 == "Mitigated Concrete", "ASR- concrete",Type2)) %>% 
  mutate(Type2 = if_else(Type2 == "Unmitigated Concrete", "ASR+ concrete",Type2)) %>% 
  select(Type, shannon, Type2) %>%
  distinct() %>%
  ggplot(aes(Type2,shannon)) +  
  geom_boxplot() + 
  geom_jitter(size = 2, width = 0.1, alpha = 0.5) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45,hjust = 1, vjust = 1)) +
  labs(x = NULL, y = "Shannon") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
    lims(y = c(2,8))

faith_precursors <- separated %>% ungroup() %>%
  #filter(Type != "Concrete") %>%
  mutate(Type2 = if_else(Type2 == "Mitigated Concrete", "ASR- concrete",Type2)) %>% 
  mutate(Type2 = if_else(Type2 == "Unmitigated Concrete", "ASR+ concrete",Type2)) %>% 
  select(Type, faith_pd, Type2) %>%
  distinct() %>%
  ggplot(aes(Type2,faith_pd)) + 
  geom_boxplot() + 
  geom_jitter(size = 2, width = 0.1, alpha = 0.5) + 
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45,hjust = 1, vjust = 1)) +
  labs(x = NULL, y = "Faith's PD") + 
  lims(y = c(0,90))

#Plot with precursors
design <- "
11223
44556
"

wrap_plots(shannon_tab, shannon, shan_precursors, faith_tab, faith, faith_precursors) + 
  plot_layout(design = design, guides = "collect",heights = c(1,1,0.3)) + 
  plot_annotation(tag_levels = 'a') + 
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/SF4_alpha_diversity_raw.pdf", 
         device = cairo_pdf, dpi = 600, width = 6.8, height = 4, scale = 1.5) +
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/SF4_alpha_diversity_raw.eps", 
         device=cairo_ps, fallback_resolution = 600, dpi = 600, width = 6.8, height = 4, scale = 1.5) +
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/SF4_alpha_diversity_raw.png", 
         type = "cairo", dpi = 600, width = 6.8, height = 4, scale = 1.5)

```


## ST4: Community composition at each taxonomic level

```{r}
library(readr)
library(phyloseq)
library(qiime2R)
library(xlsx)

metadata <- read_tsv("data/sample-metadata.tsv") %>% 
  rename(sample = "#SampleID")

table <- read_qza("data/processed/biom_tables/decontaminated/all_w_taxonomy_decontam_table.biom_table.qza")$data %>% 
  as.data.frame() %>% 
  rownames_to_column("otu") %>% 
  gather("sample","count",-"otu") %>% 
  left_join(metadata %>% select(sample,Type)) %>% 
  group_by(otu,Type) %>% 
  summarise(count = sum(count)) %>% 
  ungroup() %>% group_by(Type) %>% 
  mutate(rel_abund = count / sum(count)) %>% #make rel. abundance
  select(otu,Type,rel_abund) %>% 
  spread("Type","rel_abund") %>% 
  column_to_rownames("otu") %>% 
  otu_table(taxa_are_rows = TRUE)

metadat <- metadata %>% 
  column_to_rownames("sample") %>% sample_data()

tax <- qza_to_phyloseq(taxonomy = "data/qiime2/silva_taxonomy.qza")

decontam_ps <- phyloseq(table, tax) #%>% transform_sample_counts(function(x) x / sum(x))


levels <- c("Kingdom","Phylum","Class","Order","Family","Genus","Species")
outfile <- "g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/ST4_community_composition_all_tax_levels.xlsx"


for (level in levels){

  summarized_table <- speedyseq::tax_glom(decontam_ps,level)
  tab <- summarized_table %>% otu_table() %>% as.data.frame() %>% rownames_to_column("otu")
  tax <- summarized_table %>% tax_table() %>% .@.Data %>% as.data.frame() %>% rownames_to_column("otu") %>% left_join(tab) %>% select(-otu)
  
    
  if (level =="Kingdom"){
    completed <- level
    write.xlsx(tax %>% select(one_of(completed,colnames(tab))), 
               file= outfile, sheetName=level, row.names=FALSE)
  }
  
  if (level != "Kingdom"){
    completed <- c(completed,level)
    write.xlsx(tax %>% select(one_of(completed,colnames(tab))),
               file=outfile, sheetName=level, append=TRUE, row.names=FALSE)
  }
  
}

```


## ST6: PERMANOVA on Bray-Curtis distances

```{r message=FALSE, warning=FALSE}
library(vegan)

bc_dist <- read_qza("data/processed/biom_tables/concrete_only/all_w_taxonomy_decontam_table.biom_concrete.biom-core-metrics-results/bray_curtis_distance_matrix.qza") %>% .$data

metadata <- read_delim("data/sample-metadata.tsv", "\t", escape_double = FALSE, trim_ws = TRUE) %>% 
  as.data.frame() %>% 
  filter(Type == "Concrete") %>% 
  mutate(Temperature = scale(temp_avg),
         Months = scale(Months)) %>% 
  column_to_rownames("#SampleID") %>%
  .[rownames(as.matrix(bc_dist)),]

perm <- how(nperm = 999)
#setBlocks(perm) <- with(metadata, ASR)

set.seed(333)

##############
# sequentially
by <- "terms"
(adonis_sequential <- adonis2(bc_dist ~ Temperature + Months + ASR, data = metadata, permutations = perm, by = by))
seq_summary <- broom::tidy(adonis_sequential) %>% mutate(terms_tested = "Sequentially")

############
# Marginally
by <- "margin"
(adonis_margin <- adonis2(bc_dist ~ Temperature + Months + ASR, data = metadata, permutations = perm, by = by))
margin_summary <- broom::tidy(adonis_margin) %>% mutate(terms_tested = "Marginally")

#Calculate beta dispersion p.value for each term
beta_dist <- function(dist, term){
  anova(betadisper(dist,metadata %>% pull(paste0(term))))$`Pr(>F)`[1]
}

terms <- c("Temperature","Months","ASR")

disp <- data.frame(disp_p_val = sapply(terms,beta_dist,dist = bc_dist)) %>% 
  rownames_to_column("Coefficient")


#Make summary table

permanova_table <- bind_rows(seq_summary,margin_summary) %>% 
  mutate(Metric = "Bray-Curtis") %>%
  filter(term %in% terms) %>% 
  select(Metric, terms_tested, Coefficient = "term", R2, F = "statistic", p.value) %>% 
  left_join(disp) %>% 
  mutate(across(c("F","p.value","disp_p_val", "R2"),~round(.,3))) %>% 
  rename(`Terms Tested` = "terms_tested", `Dispersion p.value` = "disp_p_val")

knitr::kable(permanova_table)

```

## ST6: PERMANOVA on generalized (0.5) UniFrac distances, calculated before decontamination

```{r message=FALSE, warning=FALSE}
library(vegan)

gen_unidist <- read_qza("data/processed/biom_tables/concrete_only/all_w_taxonomy_table.biom_concrete.biom-core-metrics-results/generalized_05_unifrac_distance_matrix.qza") %>% .$data

metadata <- read_delim("data/sample-metadata.tsv", "\t", escape_double = FALSE, trim_ws = TRUE) %>% 
  as.data.frame() %>% 
  filter(Type == "Concrete") %>% 
  mutate(Temperature = scale(temp_avg),
         Months = scale(Months)) %>% 
  column_to_rownames("#SampleID") %>%
  .[rownames(as.matrix(gen_unidist)),]

perm <- how(nperm = 999)
#setBlocks(perm) <- with(metadata, ASR)

set.seed(333)

##############
# sequentially
by <- "terms"
(adonis_sequential <- adonis2(gen_unidist ~ Temperature + Months + ASR, data = metadata, permutations = perm, by = by))
seq_summary <- broom::tidy(adonis_sequential) %>% mutate(terms_tested = "Sequentially")

############
# Marginally
by <- "margin"
(adonis_margin <- adonis2(gen_unidist ~ Temperature + Months + ASR, data = metadata, permutations = perm, by = by))
margin_summary <- broom::tidy(adonis_margin) %>% mutate(terms_tested = "Marginally")

#Calculate beta dispersion p.value for each term
beta_dist <- function(dist, term){
  anova(betadisper(dist,metadata %>% pull(paste0(term))))$`Pr(>F)`[1]
}

terms <- c("Temperature","Months","ASR")

disp <- data.frame(disp_p_val = sapply(terms,beta_dist,dist = gen_unidist)) %>% 
  rownames_to_column("Coefficient")


#Make summary table

permanova_table <- bind_rows(seq_summary,margin_summary) %>% 
  mutate(Metric = "gUniFrac") %>%
  filter(term %in% terms) %>% 
  select(Metric, terms_tested, Coefficient = "term", R2, F = "statistic", p.value) %>% 
  left_join(disp) %>% 
  mutate(across(c("R2","F","p.value","disp_p_val"),~round(.,3)),
         p.value = if_else(p.value < 0.05, paste0(p.value,"*"),paste(p.value))) %>% 
  rename(`Terms Tested` = "terms_tested", `Dispersion p.value` = "disp_p_val")

knitr::kable(permanova_table)

gt::gt(permanova_table) %>% 
  gt::tab_style(
    style = list(
      gt::cell_text(weight = "bold")
    ),
    locations = list(
      gt::cells_column_labels(gt::everything())
    )
  )
```

## ST7: Mantel correlation tests of time and temperature with community composition

```{r}
library(tidyverse)
library(vegan)
library(qiime2R)

set.seed(333)

gUniFrac <- read_qza("data/processed/biom_tables/concrete_only/all_w_taxonomy_decontam_table.biom_concrete.biom-core-metrics-results/generalized_05_unifrac_distance_matrix.qza")$data
  
BrayCurtis <- read_qza("data/processed/biom_tables/concrete_only/all_w_taxonomy_decontam_table.biom_concrete.biom-core-metrics-results/bray_curtis_distance_matrix.qza") %>% .$data

metadata <- read_delim("data/sample-metadata.tsv", "\t", escape_double = FALSE, trim_ws = TRUE) %>% 
  as.data.frame() %>% 
  rename(Temperature = "temp_avg") %>% 
  column_to_rownames("#SampleID") %>%
  .[rownames(as.matrix(gUniFrac)),]

perm <- how(nperm = 999)
setBlocks(perm) <- with(metadata, ASR)

months_dist <- metadata %>% select(Months) %>% scale() %>% vegdist("euclid")
temp_dist <- metadata %>% select(Temperature) %>% scale() %>% vegdist("euclid")
temp_and_months_dist <- metadata %>% select(Temperature, Months) %>% scale() %>% vegdist("euclid")


mantel_res <- data.frame(dist_metric = c("gUniFrac","gUniFrac","gUniFrac","BrayCurtis","BrayCurtis","BrayCurtis"),
                         coefficients = c("Months","Temperature","Months, Temperature","Months","Temperature","Months, Temperature"),
                         coef_dist = c("months_dist","temp_dist","temp_and_months_dist","months_dist","temp_dist","temp_and_months_dist"),stringsAsFactors = F )

rho <- c()
p <- c()

for (i in 1:length(mantel_res$dist_metric)){
  
  mant <- mantel(get(mantel_res$dist_metric[i]), get(mantel_res$coef_dist[i]), permutations = 999)
  
  p <- c(p,mant$signif)
  rho <- c(rho,mant$statistic)
}

mantel_table <- mantel_res %>% bind_cols(p.val = p, rho = rho) %>% 
  mutate(across(c(p.val,rho), ~ round(.,3)), #round values
         p.val = if_else(p.val < 0.05, paste0(p.val,"*"),paste(p.val))) %>% 
  select(`Dist. metric` = "dist_metric",
         `Coefficient(s)` = "coefficients",
         `Pearson r` = "rho",
         `P-value` = "p.val")

#knitr::kable(mantel_table)
gt::gt(mantel_table) %>% 
  gt::tab_style(
    style = list(
      gt::cell_text(weight = "bold")
    ),
    locations = list(
      gt::cells_column_labels(gt::everything())
    )
  )

```

## SF6: Concrete similarity to EMP groups over time

```{r}
library(patchwork)

metadata <- read_delim("data/processed/emp_data/emp_and_tapwater_comparison_map.tsv", "\t", escape_double = FALSE, trim_ws = TRUE) %>%
  dplyr::select(sample = "#SampleID",Env,empo_2,empo_3, env_material,env_feature,env_biome,sample_scientific_name,host_common_name)

series_samples <- c("Concrete", "FlyAsh", "Powder", "Sand", "Gravel")

concrete_met <- read_tsv("data/sample-metadata.tsv")

concrete_samples <- concrete_met %>% filter(Type == "Concrete") %>% pull(`#SampleID`)

precursors <- c("FlyAsh", "Gravel", "Powder", "Sand", "Concrete")

decontam_emp_dists <- read_qza("data/processed/emp_data/core-metrics-results_emp_decontam_concrete_tapwater/general_0_5_unifrac_distance_matrix.qza")$data %>% 
  as.matrix() %>% as.data.frame() %>% 
  rownames_to_column("s1") %>% 
  gather("s2","decontam_dist",-s1) %>%
  filter(s1 != s2, s1 %in% concrete_samples) %>%
  left_join(metadata %>% dplyr::rename(s1 = "sample")) %>%
  dplyr::rename(s1_env = "Env", 
                s1_empo_2 = "empo_2", 
                s1_empo_3 = "empo_3", 
                s1_mat = "env_material", 
                s1_feature = "env_feature", 
                s1_biome = "env_biome", 
                s1_sci_name = "sample_scientific_name", 
                s1_host = "host_common_name") %>%
  left_join(metadata %>% dplyr::rename(s2 = "sample")) %>% 
  dplyr::rename(s2_env = "Env", 
                s2_empo_2 = "empo_2", 
                s2_empo_3 = "empo_3",
                s2_mat = "env_material", 
                s2_feature = "env_feature", 
                s2_biome = "env_biome", 
                s2_sci_name = "sample_scientific_name", 
                s2_host = "host_common_name") %>%
  filter(s1_empo_3 == "Concrete")


decontam_plot_data <- decontam_emp_dists %>%
  #left_join(raw_emp_dists) %>% #so that raw dists can be looked at as well
  select(s1,s2,s1_empo_3,s2_empo_3,decontam_dist) %>%
  filter(s1_empo_3 == "Concrete") %>%
  filter(!s2_empo_3 %in% precursors) %>% 
  left_join(concrete_met %>% select(s1 = "#SampleID", Months, ASR)) %>%
  group_by(Months, s2_empo_3) %>%
  summarise(mean_dist = mean(decontam_dist)) %>%
  ungroup() %>% 
  group_by(s2_empo_3) %>% 
  mutate(scaled_mean = scales::rescale(1 - mean_dist, c(0,1)), 
         similarity = 1 - mean_dist)


decontam_plot_data %>%
  left_join(concrete_met %>% select(Months, temp_avg, Collected)) %>% 
  ggplot(aes(Months, similarity, color = temp_avg)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(s2_empo_3 ~ ., ncol = 4) +
  theme_bw() +
  ggsave("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/SF6_concrete_EMPO3_sim_over_time.png", type = "cairo", width = 8, height = 6)

  
```

## SF7: Decontamination changes in concrete similarity to EMP groups

Produced in the EMP comparison .Rmd document.


## ST8: Change in prevalence of taxa over time

```{r}
library(phyloseq)
library(qiime2R)
library(tidyverse)
library(lubridate)

metadata <- read_delim("data/sample-metadata.tsv", delim = "\t") %>%
  filter(Type == "Concrete") %>%
  mutate(DoY = as.numeric(yday(mdy(Collected))),
         s_temp = scale(temp_avg),
         mean_temp = mean(temp_avg), 
         month_num = month(mdy(Collected))) %>%
  select(sample = "#SampleID", DoY,Months, mean_temp,temp_avg, Days,s_temp, month_num, Year, ASR, Type)


tree <- qza_to_phyloseq(tree = "data/qiime2/sepp_tree.qza")
tax <- qza_to_phyloseq(taxonomy = "data/qiime2/silva_taxonomy.qza")
otu_tab <- read_qza("data/processed/biom_tables/decontaminated/all_w_taxonomy_decontam_table.biom_table.qza")$data %>% 
  as.data.frame() %>% #wouldn't merge in phyloseq w/o doing this for some reason
  otu_table(taxa_are_rows = TRUE)

phylo <- merge_phyloseq(otu_tab, tax, tree)

genus_phylo <- speedyseq::tax_glom(phylo,taxrank = "Genus")


table <- otu_table(genus_phylo) %>% 
  as(Class = "matrix") %>% 
  microbiome::transform("pa") %>%
  as.data.frame()

tax <- tax_table(genus_phylo) %>% 
  as(Class = "matrix") %>% 
  as.data.frame() %>% 
  rownames_to_column("otu") %>% 
  select(-Species)


long_table <- table %>%
  rownames_to_column("otu") %>%
  gather("sample","abund",-otu) %>%
  #filter(abund > 0) %>%
  left_join(tax) %>% 
  left_join(metadata) %>%
  filter(Type == "Concrete")

tax_levels <- c("Phylum","Class","Order","Family","Genus","Species")

seasonal_taxa <- data.frame() #data.frame(Taxon = character() ,Tax_lvl = character(),model_summary = list(data.frame(Test = character())))
over_time_taxa <- data.frame()

logit_res <- data.frame()

#tax_lev <- "Phylum"
#taxa = unique_taxa[1]

for (tax_lev in tax_levels){
  
  #Run GAM for all taxa
  
  sum_phylo <- speedyseq::tax_glom(phylo,taxrank = paste(tax_lev))
  
  #Extract otu table from phyloseq obj.
  table <- otu_table(sum_phylo) %>% 
    as(Class = "matrix") %>% 
    microbiome::transform("pa") %>% #convert to presence/absence matrix
    as.data.frame()
  
  #Extract taxonomy table from pyloseq obj.
  tax <- tax_table(sum_phylo) %>% 
    as(Class = "matrix") %>% 
    as.data.frame() %>% 
    rownames_to_column("otu")
  
  #Make long/tidy table
  long_table <- table %>%
    rownames_to_column("otu") %>%
    gather("sample","abund",-otu) %>%
    #filter(abund > 0) %>%
    left_join(tax) %>% 
    left_join(metadata) %>%
    filter(Type == "Concrete") %>% 
    mutate(ASR = relevel(as.factor(ASR), "unreactive")) %>% 
    group_by(!!sym(tax_lev)) %>% 
    filter(sum(abund) >= 5) %>% #only consider taxa seen at least 5 times
    ungroup()
  
  
  unique_taxa <- as.character(unique(long_table %>% pull(tax_lev)))
  
  
  for (taxa in unique_taxa){
    
    tryCatch({
      focused_table <- long_table %>%
        filter(!!sym(tax_lev) == taxa) %>%
        group_by(Months)
      
      tax_obs <- sum(focused_table$abund)
      
      gam_taxa <- glm(abund ~ ASR*scale(Months), data = focused_table, family = "binomial")
      
      #gam_taxa <- glm(abund ~ ASR + ASR:scale(Months), data = focused_table, family = "binomial")
    
      gam_taxa_summary <- summary(gam_taxa)
      
      # if(broom::tidy(gam_taxa) %>% filter(term == "ASRreactive:scale(Months)") %>% .$p.value < 0.05 
      #    #broom::tidy(gam_taxa) %>% filter(term == "ASRreactive") %>% .$p.value < 0.05
      #    #broom::tidy(gam_taxa) %>% filter(term == "scale(Months)") %>% .$p.value < 0.05
      #    ){
      #   
        model_summary <- broom::tidy(gam_taxa)
        
        logit_res <- logit_res %>% bind_rows(
          data.frame(taxa = taxa,
          n_obs = tax_obs,
          tax_lev = tax_lev,
          num_at_tax_level = length(unique_taxa),
          aic = gam_taxa_summary$aic,
          coefs = I(list(model_summary))
        ))
     # }
    }, error=function(e){}) #if model generates error, ignore that taxa
  }
}

unnested_logit_res <- unnest(logit_res,coefs) %>% pivot_wider(id_cols =c(taxa,n_obs,tax_lev,aic,num_at_tax_level), names_from = term, values_from = estimate:p.value)


logit_res <- unnested_logit_res %>% 
  mutate(plot_name = str_remove(taxa, "D_[0-9]__")) %>% 
  mutate(log_odds_months_ASR = round(exp(1) * (`estimate_ASRreactive:scale(Months)` + estimate_ASRreactive), 2),
         `p.value_ASRreactive:scale(Months)` = signif(`p.value_ASRreactive:scale(Months)`, 2))

filt_logit_res <- logit_res %>% 
  filter(estimate_ASRreactive > 0, 
         `estimate_ASRreactive:scale(Months)` > 0,
         aic < 115)

write_tsv(logit_res,"g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/ST8_presense_absence_change.tsv")
```


## ST9:: Seasonal changes in relative abundance

```{r}
library(tidyverse)
library(qiime2R)
library(mgcv)
library(lubridate)

metadata <- read_delim("data/sample-metadata.tsv", delim = "\t") %>%
  filter(Type == "Concrete") %>%
  mutate(DoY = as.numeric(yday(mdy(Collected))),
         s_temp = scale(temp_avg),
         mean_temp = mean(temp_avg), 
         month_num = month(mdy(Collected))) %>%
  select(sample = "#SampleID", DoY,Months, mean_temp,temp_avg, Days,s_temp, month_num, Year, ASR, Type)

tree <- qza_to_phyloseq(tree = "data/qiime2/sepp_tree.qza")
tax <- qza_to_phyloseq(taxonomy = "data/qiime2/silva_taxonomy.qza")
otu_tab <- read_qza("data/processed/biom_tables/decontaminated/all_w_taxonomy_decontam_table.biom_table.qza")$data %>% 
  as.data.frame() %>% #wouldn't merge in phyloseq w/o doing this for some reason
  otu_table(taxa_are_rows = TRUE)

phylo <- merge_phyloseq(otu_tab, tax, tree)

genus_phylo <- speedyseq::tax_glom(phylo,taxrank = "Genus")

#otu table
table <- otu_table(genus_phylo) %>% 
  as(Class = "matrix") %>% 
  microbiome::transform("hell") %>% #This applies a transformation to the data, hellinger ("hell") or center-log-ratio ("clr") are recommended.
  as.data.frame()

#process taxonomy data
tax <- tax_table(genus_phylo) %>% 
  as(Class = "matrix") %>% 
  as.data.frame() %>% 
  rownames_to_column("otu") %>% 
  select(-Species)

#convert data to long form
long_table <- table %>%
  rownames_to_column("otu") %>%
  gather("sample","abund",-otu) %>%
  filter(abund > 0) %>%
  left_join(tax) %>% 
  left_join(metadata) %>%
  filter(Type == "Concrete")

#Run GAM for all taxonomic levels
#Taxa that do not appear in enough samples for the model to properly run are excluded from consideration.

tax_levels <- c("Phylum","Class","Order","Family","Genus","Species")

seasonal_taxa <- data.frame() 
over_time_taxa <- data.frame()

p.val.thresh = 1

devAskNewPage(ask = FALSE) #so no prompt for next plot


for (tax_lev in tax_levels){
  
  #Run GAM for all taxa
  
  sum_phylo <- speedyseq::tax_glom(phylo,taxrank = paste(tax_lev))
  
  
  table <- otu_table(sum_phylo) %>% 
    as(Class = "matrix") %>% 
    microbiome::transform("hell") %>%
    as.data.frame()
  
  tax <- tax_table(sum_phylo) %>% 
    as(Class = "matrix") %>% 
    as.data.frame() %>% 
    rownames_to_column("otu")
  
  long_table <- table %>%
    rownames_to_column("otu") %>%
    gather("sample","abund",-otu) %>%
    filter(abund > 0) %>%
    left_join(tax) %>% 
    left_join(metadata) %>%
    filter(Type == "Concrete")
  
  
  unique_taxa <- as.character(unique(long_table %>% pull(tax_lev)))
  
  
  for (taxa in unique_taxa){
    
    focused_table <- long_table %>%
        filter(!!sym(tax_lev) == taxa) %>%
        mutate(mean_abund = mean(abund))
    
    if (nrow(focused_table) > 10){ ##Taxa has to be observed at least 10 times
    
    tryCatch({
      
      gamm_taxa <- gamm(scale(abund) ~ s(DoY, bs = "cc",k= 4) + s(Months,bs = "cr",k=1), data = focused_table, random = list(ASR=~1))
      
      gam_taxa <- gamm_taxa$gam
      
      gam_taxa_summary <- summary(gam_taxa)
      
      if (gam_taxa_summary$s.pv[1] < p.val.thresh){
        plt <- plot(gam_taxa)[[1]]
        
        gam_month_data <- data.frame(x = plt$x, 
                             y = plt$fit,
                             se = plt$se)
        
        peak_day <- gam_month_data %>% filter(y == max(y)) %>% mutate(date = as.Date(x,origin = "2020-01-01")) %>% pull(date) %>% as.character() %>% str_remove("2020-")
        
        max_effect <- max(plt$fit)
        
        seasonal_taxa <- seasonal_taxa %>% bind_rows(data.frame(taxa, tax_lev, peak_day, max_effect, r.sq = gam_taxa_summary$r.sq, broom::tidy(gam_taxa))) 
      }
      
      if(gam_taxa_summary$s.pv[2] < p.val.thresh){
        
        plt <- plot(gam_taxa)[[2]]
        
        gam_month_data <- data.frame(x = plt$x, 
                             y = plt$fit,
                             se = plt$se)
        
        peak_month <- gam_month_data %>% filter(y == max(y)) %>% pull(x)
        
        max_effect <- max(plt$fit)
        
        over_time_taxa <- over_time_taxa %>% bind_rows(data.frame(taxa, tax_lev, peak_month,max_effect, r.sq = gam_taxa_summary$r.sq, broom::tidy(gam_taxa)))
        
      }
    }, error=function(e){}) #if model generates error, ignore that taxa
  }}
}

seasonal_wide <- seasonal_taxa %>%
  pivot_wider(names_from = term ,values_from = c(edf,ref.df,statistic,p.value)) %>% 
  write_tsv("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/ST10_long-term_relative_abundance_change.tsv")

over_time_wide <- over_time_taxa %>%
  pivot_wider(names_from = term ,values_from = c(edf,ref.df,statistic,p.value)) %>% 
  write_tsv("g:/Shared drives/Maresca Lab Shared Stuff/Concrete/Cylinder series paper/Figures/final/ST9_seasonal_relative_abundance_change.tsv")


```


## ST10: Change in relative abundance over time

Computed in the code below alongside ST9.


